Lab 8: Binomial Models in R: Answers

### 2023-03-08 Lab 8  BinomialModels/Lab8_BinomialModels.Rmd
Add files via upload

---
title: 'Lab 8: Binomiaal Models in R'
author: "Eric Parajon"
header-includes:
- \usepackage{enumerate}
- \usepackage{graphicx}
output:
  html_document:
    df_print: paged
---

Today's lab has been adapted from code by Chelsea Estancona, Isabel Laterzo, and Simon Hoellerbauer.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

@@

#Load in packages
pacman::p_load(tidyverse,
               rio,#for importing data
               arm)#inverse link function

@@

```{r}
ggplot(colombia, aes(x = as.factor(AUC))) +
  geom_bar() +
  labs(x = 'AUC Presence') +
  theme_bw()

@@

```{r}
## calculate value of the linear predictor for this variable profile
linpred <- as.numeric(with(colombia, coef(m1)[1] +
                             coef(m1)[2] * mean(tribut_2_log, na.rm = T) +
                             coef(m1)[3] * mean(tribut_2_log, na.rm = T)^2 +
                             coef(m1)[4] * mean(pop_attacks, na.rm = T) +
                             coef(m1)[5] * mean(disbogota, na.rm = T) +
                             coef(m1)[6] * (1) + #could set this to its mode
                             coef(m1)[7] * mean(pop_log, na.rm = T)))

#let's take a look!
invlogit #no parentheses from the arm package

## invlogit lets us transform our log-odds to the range (0,1) - now interpretable
## as a standard probability
invlogit(linpred)
```
We can also do this using the 'predict' function:

```{r}
## create a 'new data' frame: we want to predict the probability for a hypothetical situation.
new_data <- with(colombia, data.frame(tribut_2_log = mean(tribut_2_log, na.rm = T),
                                      tribut_2_log_sq = mean(tribut_2_log, na.rm = T)^2,
                                      pop_attacks = mean(pop_attacks, na.rm = T),
                                      disbogota = mean(disbogota, na.rm = T),
                                      drugs = c(0, 1),
                                      pop_log = mean(pop_log, na.rm = T)))
## note above that I include both possible values for drugs. the 'data.frame()'
## command is flexible and will intelligently duplicate the other values

@@

```{r}
min_tax <- min(colombia$tribut_2_log, na.rm = T)
max_tax <- max(colombia$tribut_2_log, na.rm = T)
min_tax_2 <- min(colombia$tribut_2_log_sq, na.rm = T)
max_tax_2 <- max(colombia$tribut_2_log_sq, na.rm = T)

new_data_2 <- with(colombia,
                 data.frame(tribut_2_log = seq(from = min_tax, to = max_tax,
                                               length.out = 100),
                            tribut_2_log_sq = seq(from = min_tax_2, to = max_tax_2,
                                                  length.out = 100),
                            pop_attacks = mean(pop_attacks, na.rm = T),
                            disbogota = mean(disbogota, na.rm = T),
                            drugs = 1, pop_log = mean(pop_log, na.rm = T)))

new_data_3 <- with(colombia,
                 data.frame(tribut_2_log = seq(from = min_tax, to = max_tax,
                                               length.out = 100),
                            tribut_2_log_sq = seq(from = min_tax_2, to = max_tax_2,
                                                  length.out = 100),
                            pop_attacks = mean(pop_attacks, na.rm = T),
                            disbogota = mean(disbogota, na.rm = T),
                            drugs = 0, pop_log = mean(pop_log, na.rm = T)))

@@

## plot predicted probabilities
ggplot(data = plot_pred1, aes(x = tax, y = Pred_Probs, color = link)) +
  geom_line() +
  scale_color_discrete(name = 'Link Function') +
  labs(x = 'Tax Revenue', y = 'Predicted Probability') +
  theme_bw()
```

@@

## create a vector that is a sequence of 100 points, evenly spaced between the
## lowest and highest observed values of tax income in the data set. we have to
## do this for the (tax^2) variable, too!
tax_min <- with(colombia, min(tribut_2_log, na.rm = T))
tax_max <- with(colombia, max(tribut_2_log, na.rm = T))
tax_vec <- seq(from = tax_min, to = tax_max, length = 100)
trans_vec <- (tax_vec^2)
## these will be the x values over which we can plot the predicted probability.
## we're interested in how the predicted probability of AUC presence varies over
## the municipalities' values of tax revenue

## create a hypothetical independent variable as before: mean attacks on population,
## mean distance from the capitol, drug crops set to 1, and the average population.
hyp <- with(colombia, data.frame(intercept = 1, 
                                 tribut_2_log = tax_vec,
                                 tribut_2_log_sq = trans_vec,
                                 pop_attacks = mean(pop_attacks, na.rm = T),
                                 disbogota = mean(disbogota, na.rm = T),
                                 drugs = 1,
                                 pop_log = mean(pop_log, na.rm = T)))

@@

  pp_sims[i, ] <- invlogit(as.matrix(hyp) %*% beta_sims[i, ])

@@

#we could also do this with less code:
pp_sims2 <- invlogit(as.matrix(hyp) %*% t(beta_sims))

@@

    pp <- invlogit(temporary %*% t(beta_sims))

@@

sim_pp <- apply(sim_betas, 1, function(x) invlogit(as.matrix(sim_dat) %*% x))

@@

    pp <- invlogit(temporary %*% t(sim_betas_m1))

>>>>

---
title: 'Lab 8: Binomial Models in R'
author: "Eric Parajon"
header-includes:
- \usepackage{enumerate}
- \usepackage{graphicx}
output:
  html_document:
    df_print: paged
---

Today's lab has been adapted from code by Chelsea Estancona, Isabel Laterzo, and Simon Hoellerbauer.
```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

@@

#Load in packages
pacman::p_load(tidyverse,
               rio)#for importing data

@@

```{r}
ggplot(colombia, aes(x = as.factor(AUC))) +
  geom_bar() +
  labs(x = 'Autodefensas Unidas de Colombia Presence') +
  theme_bw()

@@

```{r}
## calculate value of the linear predictor for this variable profile
linpred <- as.numeric(with(colombia, coef(m1)[1] +
                             coef(m1)[2] * mean(tribut_2_log, na.rm = TRUE) +
                             coef(m1)[3] * mean(tribut_2_log, na.rm = TRUE)^2 +
                             coef(m1)[4] * mean(pop_attacks, na.rm = TRUE) +
                             coef(m1)[5] * mean(disbogota, na.rm = TRUE) +
                             coef(m1)[6] * (1) + #could set this to its mode
                             coef(m1)[7] * mean(pop_log, na.rm = TRUE)))

#let's take a look!
stats::plogis

## plogis lets us transform our log-odds to the range (0,1) - now interpretable
## as a standard probability

plogis(linpred)
```
We can also do this using the 'predict' function:

```{r}
## create a 'new data' frame: we want to predict the probability for a hypothetical situation.
new_data <- with(colombia, data.frame(tribut_2_log = mean(tribut_2_log, na.rm = TRUE),
                                      tribut_2_log_sq = mean(tribut_2_log, na.rm = TRUE)^2,
                                      pop_attacks = mean(pop_attacks, na.rm = TRUE),
                                      disbogota = mean(disbogota, na.rm = TRUE),
                                      drugs = c(0, 1),
                                      pop_log = mean(pop_log, na.rm = TRUE)))
## note above that I include both possible values for drugs. the 'data.frame()'
## command is flexible and will intelligently duplicate the other values

@@

```{r}
min_tax <- min(colombia$tribut_2_log, na.rm = TRUE)
max_tax <- max(colombia$tribut_2_log, na.rm = TRUE)
min_tax_2 <- min(colombia$tribut_2_log_sq, na.rm = TRUE)
max_tax_2 <- max(colombia$tribut_2_log_sq, na.rm = TRUE)

new_data_2 <- with(colombia,
                 data.frame(tribut_2_log = seq(from = min_tax, to = max_tax,
                                               length.out = 100),
                            tribut_2_log_sq = seq(from = min_tax_2, to = max_tax_2,
                                                  length.out = 100),
                            pop_attacks = mean(pop_attacks, na.rm = TRUE),
                            disbogota = mean(disbogota, na.rm = TRUE),
                            drugs = 1, pop_log = mean(pop_log, na.rm = TRUE)))

new_data_3 <- with(colombia,
                 data.frame(tribut_2_log = seq(from = min_tax, to = max_tax,
                                               length.out = 100),
                            tribut_2_log_sq = seq(from = min_tax_2, to = max_tax_2,
                                                  length.out = 100),
                            pop_attacks = mean(pop_attacks, na.rm = TRUE),
                            disbogota = mean(disbogota, na.rm = TRUE),
                            drugs = 0, pop_log = mean(pop_log, na.rm = TRUE)))

@@

## plot predicted probabilities
ggplot(data = plot_pred1, aes(x = tax, y = Pred_Probs, color = link)) +
  geom_line() +
  scale_color_discrete(name = 'Link Function') +
  labs(x = 'Tax Revenue', y = 'Predicted Probability') +
  theme_bw()+
  geom_rug(data = colombia, aes(x = tribut_2_log), 
           size = 1, inherit.aes = F) #adding rug plot what is this showing us? Why is there a divergence?

@@

## create a vector that is a sequence of 100 points, evenly spaced between the
## lowest and highest observed values of tax income in the data set. we have to
## do this for the (tax^2) variable, too!
tax_min <- with(colombia, min(tribut_2_log, na.rm = TRUE))
tax_max <- with(colombia, max(tribut_2_log, na.rm = TRUE))
tax_vec <- seq(from = tax_min, to = tax_max, length = 100)
trans_vec <- (tax_vec^2)
## these will be the x values over which we can plot the predicted probability.
## we're interested in how the predicted probability of AUC presence varies over
## the municipalities' values of tax revenue

## create a hypothetical independent variable as before: mean attacks on population,
## mean distance from the capitol, drug crops set to 1, and the average population.
hyp <- with(colombia, data.frame(intercept = 1, 
                                 tribut_2_log = tax_vec,
                                 tribut_2_log_sq = trans_vec,
                                 pop_attacks = mean(pop_attacks, na.rm = TRUE),
                                 disbogota = mean(disbogota, na.rm = TRUE),
                                 drugs = 1,
                                 pop_log = mean(pop_log, na.rm = TRUE)))

@@

  pp_sims[i, ] <- plogis(as.matrix(hyp) %*% beta_sims[i, ])

@@

#we could also do this with less code:
pp_sims2 <- plogis(as.matrix(hyp) %*% t(beta_sims))

@@

    pp <- plogis(temporary %*% t(beta_sims))

@@

sim_pp <- apply(sim_betas, 1, function(x) plogis(as.matrix(sim_dat) %*% x))

@@

    pp <- plogis(temporary %*% t(sim_betas_m1))

### 2023-03-08 Lab 8  BinomialModels/Lab8_BinomialModels_AnswerKey.Rmd
Add files via upload

---
title: 'Lab 8: Binomiaal Models in R: Answers'
author: 'TA: Eric Parajon'
output:
  html_document:
    df_print: paged
---

Today's lab has been adapted from code by Chelsea Estancona, Isabel Laterzo, and Simon Hoellerbauer.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

@@

#Load in packages
pacman::p_load(tidyverse,
               rio,
               arm)#for importing data
```

@@

```{r}
ggplot(colombia, aes(x = as.factor(AUC))) +
  geom_bar() +
  labs(x = 'AUC Presence') +
  theme_bw()
```

@@

```{r}
## calculate value of the linear predictor for this variable profile
linpred <- as.numeric(with(colombia, coef(m1)[1] +
                             coef(m1)[2] * mean(tribut_2_log, na.rm = T) +
                             coef(m1)[3] * mean(tribut_2_log, na.rm = T)^2 +
                             coef(m1)[4] * mean(pop_attacks, na.rm = T) +
                             coef(m1)[5] * mean(disbogota, na.rm = T) +
                             coef(m1)[6] * (1) + #could set this to its mode
                             coef(m1)[7] * mean(pop_log, na.rm = T)))

#let's take a look!
arm::invlogit #no parentheses from the arm package

## invlogit lets us transform our log-odds to the range (0,1) - now interpretable
## as a standard probability
invlogit(linpred)
```
We can also do this using the 'predict' function:

```{r}
## create a 'new data' frame: we want to predict the probability for a hypothetical situation.
new_data <- with(colombia, data.frame(tribut_2_log = mean(tribut_2_log, na.rm = T),
                                      tribut_2_log_sq = mean(tribut_2_log, na.rm = T)^2,
                                      pop_attacks = mean(pop_attacks, na.rm = T),
                                      disbogota = mean(disbogota, na.rm = T),
                                      drugs = c(0, 1),
                                      pop_log = mean(pop_log, na.rm = T)))
## note above that I include both possible values for drugs. the 'data.frame()'
## command is flexible and will intelligently duplicate the other values

@@

```{r}
min_tax <- min(colombia$tribut_2_log, na.rm = T)
max_tax <- max(colombia$tribut_2_log, na.rm = T)
min_tax_2 <- min(colombia$tribut_2_log_sq, na.rm = T)
max_tax_2 <- max(colombia$tribut_2_log_sq, na.rm = T)

new_data_2 <- with(colombia,
                 data.frame(tribut_2_log = seq(from = min_tax, to = max_tax,
                                               length.out = 100),
                            tribut_2_log_sq = seq(from = min_tax_2, to = max_tax_2,
                                                  length.out = 100),
                            pop_attacks = mean(pop_attacks, na.rm = T),
                            disbogota = mean(disbogota, na.rm = T),
                            drugs = 1, pop_log = mean(pop_log, na.rm = T)))

new_data_3 <- with(colombia,
                 data.frame(tribut_2_log = seq(from = min_tax, to = max_tax,
                                               length.out = 100),
                            tribut_2_log_sq = seq(from = min_tax_2, to = max_tax_2,
                                                  length.out = 100),
                            pop_attacks = mean(pop_attacks, na.rm = T),
                            disbogota = mean(disbogota, na.rm = T),
                            drugs = 0, pop_log = mean(pop_log, na.rm = T)))

@@

#we can calculate the logit predicted probs by hand by doing 1/(1 + exp(-linpred))
#How could we calculate this for probit?

@@

```{r, cache = TRUE}
## create a 'temporary' dataset - bind a 1 for the intercept and omit missingness
temporary <- model.matrix( ~ tribut_2_log + tribut_2_log_sq + pop_attacks + 
                             disbogota + drugs + pop_log, data = colombia)

@@

```{r}
pp <- map_dfc(income_vec, function(j) {
    temporary[, "income"] <- j

    pp <- invlogit(temporary %*% t(sim_betas_m1))

@@

head(temporary)
```
2. We then matrix multiply this new design matrix, with all actual values of our predictor of interest replaced by one hypothetical value of our predictor interest and with all the other values held at their observed values, by our simulated betas. This creates a matrix, which has as many rows as the rows of our data, and as many columns as there are samples of our betas.  
```{r}
dim(invlogit(temporary %*% t(sim_betas_m1)))
```
3. We then take the mean across the columns (`margins = 2`) because different columns represent different predicted probabilities, for each of the observations in our temporary design matrix, when we use different betas and when our predictor of interest has been replaced with one value. This means that we then have a mean predicted probability for each different set of betas when our predictor of interest is at a certain value.

>>>>

---
title: 'Lab 8: Binomial Models in R: Answers'
author: 'TA: Eric Parajon'
output:
  html_document:
    df_print: paged
---

Today's lab has been adapted from code by Chelsea Estancona, Isabel Laterzo, and Simon Hoellerbauer.
```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE)

@@

#Load in packages
pacman::p_load(tidyverse,
               rio)#for importing data
```

@@

```{r}
ggplot(colombia, aes(x = as.factor(AUC))) +
  geom_bar() +
  labs(x = 'Autodefensas Unidas de Colombia Presence') +
  theme_bw()
```

@@

```{r}
## calculate value of the linear predictor for this variable profile
linpred <- as.numeric(with(colombia, coef(m1)[1] +
                             coef(m1)[2] * mean(tribut_2_log, na.rm = TRUE) +
                             coef(m1)[3] * mean(tribut_2_log, na.rm = TRUE)^2 +
                             coef(m1)[4] * mean(pop_attacks, na.rm = TRUE) +
                             coef(m1)[5] * mean(disbogota, na.rm = TRUE) +
                             coef(m1)[6] * (1) + #could set this to its mode
                             coef(m1)[7] * mean(pop_log, na.rm = TRUE)))

#let's take a look!
stats::plogis

## plogis lets us transform our log-odds to the range (0,1) - now interpretable
## as a standard probability

plogis(linpred)
```
We can also do this using the 'predict' function:

```{r}
## create a 'new data' frame: we want to predict the probability for a hypothetical situation.
new_data <- with(colombia, data.frame(tribut_2_log = mean(tribut_2_log, na.rm = TRUE),
                                      tribut_2_log_sq = mean(tribut_2_log, na.rm = TRUE)^2,
                                      pop_attacks = mean(pop_attacks, na.rm = TRUE),
                                      disbogota = mean(disbogota, na.rm = TRUE),
                                      drugs = c(0, 1),
                                      pop_log = mean(pop_log, na.rm = TRUE)))
## note above that I include both possible values for drugs. the 'data.frame()'
## command is flexible and will intelligently duplicate the other values

@@

```{r,results="hide"}
min_tax <- min(colombia$tribut_2_log, na.rm = TRUE)
max_tax <- max(colombia$tribut_2_log, na.rm = TRUE)
min_tax_2 <- min(colombia$tribut_2_log_sq, na.rm = TRUE)
max_tax_2 <- max(colombia$tribut_2_log_sq, na.rm = TRUE)

new_data_2 <- with(colombia,
                 data.frame(tribut_2_log = seq(from = min_tax, to = max_tax,
                                               length.out = 100),
                            tribut_2_log_sq = seq(from = min_tax_2, to = max_tax_2,
                                                  length.out = 100),
                            pop_attacks = mean(pop_attacks, na.rm = TRUE),
                            disbogota = mean(disbogota, na.rm = TRUE),
                            drugs = 1, pop_log = mean(pop_log, na.rm = TRUE)))

new_data_3 <- with(colombia,
                 data.frame(tribut_2_log = seq(from = min_tax, to = max_tax,
                                               length.out = 100),
                            tribut_2_log_sq = seq(from = min_tax_2, to = max_tax_2,
                                                  length.out = 100),
                            pop_attacks = mean(pop_attacks, na.rm = TRUE),
                            disbogota = mean(disbogota, na.rm = TRUE),
                            drugs = 0, pop_log = mean(pop_log, na.rm = TRUE)))

@@

```{r}
#we can calculate the logit predicted probs by hand by doing 1/(1 + exp(-linpred))
#How could we calculate this for probit?

@@

```{r,message=FALSE, results='hide'}
## create a 'temporary' dataset - bind a 1 for the intercept and omit missingness
temporary <- model.matrix( ~ tribut_2_log + tribut_2_log_sq + pop_attacks + 
                             disbogota + drugs + pop_log, data = colombia)

@@

```{r,message=FALSE, results='hide'}
pp <- map_dfc(income_vec, function(j) {
    temporary[, "income"] <- j

    pp <- plogis(temporary %*% t(sim_betas_m1))

@@

head(temporary)
```
2. We then matrix multiply this new design matrix, with all actual values of our predictor of interest replaced by one hypothetical value of our predictor interest and with all the other values held at their observed values, by our simulated betas. This creates a matrix, which has as many rows as the rows of our data, and as many columns as there are samples of our betas.  
```{r}
dim(plogis(temporary %*% t(sim_betas_m1)))
```
3. We then take the mean across the columns (`margins = 2`) because different columns represent different predicted probabilities, for each of the observations in our temporary design matrix, when we use different betas and when our predictor of interest has been replaced with one value. This means that we then have a mean predicted probability for each different set of betas when our predictor of interest is at a certain value.
